---
title: "Data clean"
author: "Michael Robinson"
date: "2023-10-11"
output:
  html_document: default
  pdf_document: default
---

# Dataset 1
# Invoices With Merged Categories and Merged Amounts
Because a single transaction (identified with an order id ) has multiple items purchased, who ever captured this data decided to create a single row for each order, thereby lumping the different items purchased and the amounts together into 2 fields respectively.

The better thing to do is to let each item purchased be on a single row with the amount. It is better to repeat the Order IDs on different rows than lumping up amounts in a single cell. We would be analyzing items bought and amounts a lot, we need them separated into rows.

```{r cars}
library(dplyr)
library(tidyr)


invoice_data <- "https://raw.githubusercontent.com/MRobinson112/data_607_project2/main/Dirty_invoice.csv"

data <- read.csv(invoice_data, header = TRUE, stringsAsFactors = FALSE)

print(data)

# Split the Category and Amount columns by "|"
data_split <- data %>%
  separate_rows(Category, Amount, sep = "\\|")

# View the cleaned data in the desired format
print(data_split)

``` 
#Dataset 2
# Medicine Data With Combined Quantity and Measure
Going by clean data rules, you should have every field/column represent unique things. So split the combined Quantity and Measure on this data into separate columns/fields. When you are done, your Quantity column should sum up to 17,600.00. You will find this total on the clean worksheet once you download.

```{r}
# Load the tidyr package
library(tidyr)
library(readr)

#
Med_data <- "https://raw.githubusercontent.com/MRobinson112/data_607_project2/main/med_data.csv"

datam <- read.csv(Med_data, header = TRUE, stringsAsFactors = FALSE)

datam

df <- separate(datam, Quantity, into = c("Quantity", "Measure"), sep = "(?<=\\d)(?=[A-Za-z])", convert = TRUE)

# convert = TRUE will try to automatically convert quantity to numeric

# Print the resulting data frame
print(df)

``` 
# Dataset #3 - a work order process at a manufacturing site
Question: What is the distribution of work order status by Building and Type?
This was taken from slack uploades by Sam
```{r}

library(dplyr)
library(tidyr)
library(tidyxl)
library(ggplot2)


Order_data <- readxl::read_excel("~/Downloads/Example of a work order process at a manufacturing site.xlsx", skip = 1) 

head(Order_data)
Order_data <- Order_data %>% filter(!is.na(Building) | !is.na(Type))

Order_data <- Order_data %>% fill(Building, Type)

tidy_data <- Order_data %>%
  pivot_longer(cols = -c(Building, Type), names_to = "Work_Order_Status", values_to = "Value")

head(tidy_data) 

# a bar plot to show the distribution of work order status by Building and Type
ggplot(tidy_data, aes(x = Building, fill = Work_Order_Status)) +
  geom_bar(position = "fill") +
  facet_wrap(~ Type) +
  labs(title = "Distribution of Work Order Status by Building and Type",
       y = "Proportion",
       x = "Building") +
  theme_minimal()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
